---
title: World War II
author: Dimitri Lozeve
output:	html_document
fig_height: 7
fig_width: 8
---


## Libraries

```{r}
library(tidyverse)
library(lubridate)
```

Package `lubridate` is used for date and time management.


## Data import

I used the column types from data.world. I also parse the "Mission
date" column using `lubridate`.

```{r results="hide"}
bombings <- read_csv("THOR_WWII_DATA_CLEAN.csv",col_types="iicccccccciicccddccccccddddcdddcdddccdddccddcccdcccccccccccccc")
bombings$MSNDATE <- mdy(bombings$MSNDATE)
glimpse(bombings)
```

## Total tons of explosive by country

List of all countries targeted during WWII by the US:

```{r}
levels(as.factor(bombings$TGT_COUNTRY))
```

To plot the total amount of explosives dropped on each country, I
build a new table, which I sort and fix the order using factors.

```{r}
sum_by_country <- bombings %>% group_by(TGT_COUNTRY) %>% summarise(TOTAL_TONS= sum(TOTAL_TONS, na.rm=TRUE))

sum_by_country <- sum_by_country[order(sum_by_country$TOTAL_TONS, decreasing=TRUE),]
sum_by_country$TGT_COUNTRY <- factor(sum_by_country$TGT_COUNTRY, levels = sum_by_country$TGT_COUNTRY)
sum_by_country
```

I then plot a bar plot for the first 10 countries:

```{r}
ggplot(top_n(sum_by_country, 10, TOTAL_TONS), aes(x=TGT_COUNTRY, y=TOTAL_TONS)) +
    geom_bar(stat="identity") +
    #scale_y_log10() + ## for a log plot
    theme(axis.text.x = element_text(angle = 50, hjust = 1))
```

## Total tons of explosive by month

In the same manner as before, I build a new table with a column for the month.

```{r}
sum_by_month <- bombings %>% select(MSNDATE, TOTAL_TONS)
sum_by_month$MONTH <- ymd(format(bombings$MSNDATE, "%Y-%m-1"))
```

```{r}
ggplot(sum_by_month, aes(x=MONTH, y=TOTAL_TONS)) +
    stat_summary(fun.y=sum, geom="bar") +
    scale_x_date(date_labels="%Y", date_breaks="1 year", date_minor_breaks="1 month")
```

## Map of bombings in Europe

```{r}
bombings_locations<- bombings %>%
    group_by(LONGITUDE, LATITUDE) %>%
    summarise(TOTAL_TONS = sum(TOTAL_TONS))

bombings_locations <- na.omit(bombings_locations)
#bombings_locations <- left_join(bombings_locations, select(bombings, LONGITUDE, LATITUDE, TGT_LOCATION), by=c("LONGITUDE", "LATITUDE"))
```

```{r}
library(leaflet)
```

```{r fig.height=7, fig.width=10 }
m <- leaflet(na.omit(bombings_locations)) %>%
    setView(lng = 20, lat = 20, zoom = 3) %>%
    addTiles(group = "OpenStreetMap") %>%
    addProviderTiles("Stamen.Toner", group = "Toner") %>%
    addProviderTiles("Stamen.TonerLite", group = "TonerLite") %>%
    addProviderTiles("Esri.WorldTopoMap", group = "Topographical") %>%
    addProviderTiles("Esri.NatGeoWorldMap", group = "NatGeo") %>%
    addProviderTiles("Esri.WorldImagery", group = "Satellite") %>%
    addCircles(lng = ~LONGITUDE, lat = ~LATITUDE,
               radius = ~sqrt(TOTAL_TONS) * 200,
               fillColor = "red", color = "red",
               group = "World War II") %>%
    addLayersControl(
        baseGroups = c("OpenStreetMap", "Toner", "TonerLite", "Topographical", "NatGeo", "Satellite"),
        overlayGroups = c("World War II"),
        options = layersControlOptions(collapsed = FALSE)) %>%
    showGroup("TonerLite")
m
```
